Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-06-28 09_Article/Simon_2021_Bone_Draft_V01.tex
Second article co-authors revision

@@

\subsection{Statistics}
The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI groups. As the groups do not have similar distributions of age and sex, a matching was performed by selecting similar individuals leading to identical mean and median age as well as identical gender distribution. For each parameter, the median value between the six ROIs from the same individual was computed. The median was preferred over the mean because it is less influenced by outliers. Normality of the distribution was assessed with QQ plot and Shapiro-Wilk test. CV had to be log-transformed to meet normal distribution assumption. Then, equal variances was assessed using Bartlett test or Brown-Forsythe test according to the normality distribution of the data. According to the normality and equal variances assumptions, t-test, Mann-Whitney test or a non-parametric permutation test was performed. The general significance level was set to 95\% for all tests. Confidence intervals was computed for t-tested variables to confirm a difference in both groups means. As Mann-Whitney tests are performed on the median, only the corresponding p-value is presented. Finally, non-parametric permutation tests are less powerful but give an empirical 95\% exclusion range and a p-value. If the difference in means belong to this exclusion range, it can be stated that group means are different with 95\% certainty.

@@

\subsection{Morphological Analysis}
The results of morphological analysis are summarized in Table \ref{Table2}. In the present study, the individual matching allowed us to have similar group distributions with 17 females and 11 males in each group. The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI. BV/TV of healthy individuals is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01. Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001. The trabecular thickness do not show significant differences between groups with a p value of 0.2. On the other hand, trabecular separation is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$). Trabecular separation SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$). SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively. Finally, the log transformation of coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333].\\

In Table \ref{Table2}, absolute values and p values of test statistics are compared with literature. Regarding age, the present population is slightly younger than in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} but stay in a similar range. In a general way p values of the three other studies (\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018}) show significant differences for BV/TV, Tb.N., Tb.Sp., and Tb.Sp.SD and non significant differences for Tb.Th. Absolute values of BV/TV, TB.Th., Tb.Sp., and Tb.Sp.SD are higher in the present study compared to literature. On the other hand, Tb.N. appears to be higher in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} than in the present one.

@@

\subsection{Linear Regression with Original Data Sets}
Figure \ref{02_GeneralRegression} shows the results of the regression analysis for the complete data sets of each group separately. The x axis represent the values of observed tensors from \si{\micro}FE simulations. The y axis is the predicted values using the Zysset-Curnier model and the parameters obtained after performing the regression with linear mixed-effect model. The fitted line is represented by the dashed line, indicating the theoretical perfect correlation. $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components. For the healthy group, see Figure \ref{02_Healthy}, the fit is performed on 720 ROIs leading to 8640 points. The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%. The regression analysis for the OI group led to 3528 points, a $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%. It can be noticed that, as the values of the observed tensor decreases, points tends to be further apart of the diagonal (dashed line). Moreover, the points coming from the tensors with lowest values are only above the diagonal. One last observation which should be noted is the range covered by the observed tensors. The range of the OI group is wider than the healthy group and ROIs with lower BV/TV present further low stiffness values. The values of these ROIs tensors components tends to be overestimated by the fit.\\

@@

\subsection{Filtering}
The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}. The OI data reach higher values of CV and stay at lower values of BV/TV than healthy data. Generally, the CV tends to increase with decreasing BV/TV. The Spearman coefficient is shown above the plot as value [95\% CI]. Its value is negative and strictly different from zero. Finally, the CV threshold value used to filter the data is represented by the dashed line. It can be observed that a relatively important part of OI data will be filtered out. On the other hand, relatively few healthy data will be removed by the filtering. Examples of extreme ROIs in terms of CV and BV/TV are shown in Appendix \ref{A2}.\\

@@

The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}. After filtering, the healthy group was reduced to 119 individuals and 603 ROIs. This led to 7236 points for the regression. Results show similar values to the complete data set, namely a $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\%, see Figure \ref{04_Healthy}. The OI group lost more individuals after filtering leading to 38 people and 115 ROIs. Regression is performed on 1380 points then. Figure \ref{04_OI} presents results with a $R^2_{adj}$ rounded to 0.95 and a NE of 17\% $\pm$ 8\%.\\

@@

\subsection{BV/TV and DA Matching}
Regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}. The columns show which data set was used, the fives parameters of the Zysset-Curnier model and the assessment of fit quality. Grouping healthy and OI data together for regression lead to a $k$ of 1.91 and a $l$ of 0.95. Regression result shows a $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%. Second and last rows show regression results using separated data sets and imposing the exponents $k$ and $l$. OI values are higher than healthy one. The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively. The ANCOVA performed to quantify the group statistical significance shows a p value of 0.7.\\

@@

Table \ref{Table4} shows results obtained compared to literature. \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs. Data sets of \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} show BV/TV ranges slightly higher than in the present study and the one of \citeauthor{Gross2013} \cite{Gross2013}. On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}. Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set than for the other studies \cite{Gross2013,Panyasantisuk2015}.\\

@@

The analysis of tBMD in relation to BV/TV is shown in Figure \ref{02_tBMD}. The t-test performed on the tBMD distributions led to a p-value <0.0001 and a 95\% CI of [-35,-13]. The linear regression performed using the linear mixed-effects model without BV/TV and group interaction provided a slope of 223 [153,292] (value [95\% CI]). The intercept was 534 [518,550] and the group variable led to a value of 14 [7,20]. The prediction using the fixed effects only led to a $R^2$ of 0.23 and a SE of 33. Then, using the linear mixed-effect model with the interaction regressor (BV/TV x Group), this added variable presented a value of 3 [-66,72] and a p-value of 0.94.

@@

The age of matched individuals lies in the spectrum of the others studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} which allow one to compare the morphological values. The main explanation for differences between the absolute values of the present study compared to the others lies in the imaging system. \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed first generation Xtreme CT scans with a voxel size of 82 \si{\micro}m and the present study used a second generation scanner, XCTII, with a voxel size of 61 \si{\micro}m. The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types. They showed that BV/TV, Tb.Th., and Tb.Sp. is higher in second generation Xtreme CT and, on the other hand, Tb.N. is higher in first generation Xtreme CT. These results give confidence to the observed values. Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length. This conditions the Tb.N. and Tb.Sp. as they depend on the ROI size. Moreover, conditions imposed for ROI random selection can lead to further biased values, especially for OI patients, as it can not be empty of trabecular bone. The CV presenting the stronger significant difference between groups even given the low sample size show that heterogeneity is a main difference in OI patients compared to healthy individuals. Finally, the significant differences observed in BV/TV and DA even with age and gender matched individuals justify the choice of a variable matching for fabric-elasticity relationships analysis.\\

Regression performed on original data sets shows $R^2_{adj}$ and NE in the expected range for the healthy group. Components of the stiffness tensors are distributed to both sides of the diagonal. On the other hand, regression of the OI data set presents lower $R^2_{adj}$ than such fit reach usually. The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation. These differences come from the ROIs presenting a low stiffness. It can be seen on the regression plot (Figure \ref{02_OI}) that when the stiffness term decreases lower than $10^0$ MPa, the fit tends to overestimate the stiffness. This is due to the fact that ROI stiffness is highly impacted by BV/TV. Some low BV/TV ROIs do not have every side of the cube connected by bone leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}. Trying to homogenize such ROI can lead to error of multiple order of magnitude, as observed on the plot. Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}. An alternative of CV filtering could be to compute the area ratio filled by bone on each of the six faces of the ROI to assess the ROI heterogeneity.\\

Figure \ref{02_CV_BVTV} presenting the CV in relation to BV/TV shows that there is a tendency of CV to increase with a decreasing BV/TV. Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly becomes highly heterogeneous. A simple assumption about this relation is that it could be monotonic. Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation. As some higher BV/TV ROIs still present high CV, imposing a fixed threshold make sense for subsequent homogenization. However, the value of this threshold actually results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\

The fits performed on filtered data sets present direct effect of filtering. For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) is negligible. On the other hand, NE presents an improved value decreasing by 2\%. These results are due to the filtering of points further away from the diagonal (better NE) and some points close to the diagonal leading to a smaller number of points ($R^2_{adj}$). For the OI group, filtering leads to a important improvement in the fit. Results become similar to the healthy group in terms of $R^2_{adj}$, NE, and the range of stiffness values. These results give confidence to the filtering procedure and are a first step to the hypothesis of there being no difference in fabric-elasticity relationships between healthy and OI trabecular bone.\\

After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for individual filtered data sets. This allows one to give values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m. Imposing these values to perform the fit on individual matched data sets allows us to highlight differences, if any, between healthy and OI trabecular bone. The relatively low differences for $\lambda_0'$ and $\mu_0$ once again provide arguments for similar relationships between the two groups. For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but 95\% CI intervals still show a common range. Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group shown a p value far above the 5\% significance level. With this statistical non-significance of the group and low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to suppose differences in fabric-elasticity relationships between healthy and OI trabecular bone. In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity. Nevertheless, the error created by such ROIs are negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\

Imposing $k$ and $l$ allows one to estimate the effect of different image resolutions.  \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m. \citeauthor{Gross2013}\cite{Gross2013} shown that different anatomical locations lead to only slight differences. Comparing regression of filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution. Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones. Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges stay overlapped. As for filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies. Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants by resolution. The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explain this increase of DA. Lower stiffness constants are obtained because the coarser structure resulting from XCTII can not be as optimized as the fine detailed structure obtained by \si{\micro}CT. Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimized morphology of trabecular bone with a high fidelity. By decreasing the scan spatial resolution, the scanned structure becomes more bulky. Performing a fit on this less optimized structure lead to the observed lower stiffness constants. Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality. Nevertheless, $l$ stands in the same range as for the two other studies meaning the relative weight of DA remains constant. On the other hand, the higher $k$ highlight an increased relative weight of BV/TV.\\

@@

The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans. Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics. Effectively as those patients are in wheelchair, it could be interesting to analyze the impact of this condition on the weight bearing tibia. The ROI homogeneity has an important impact on the analysis quality. As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting. More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible. Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution. Another limitation is that the scans were performed on different devices which were not cross-calibrated. However, as they are the same model, it is expected to have a minor impact.\\

In conclusions, the sample analyzed in the this study have similar morphology to data reported in the literature. We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough. \citeauthor{Indermaur2021}\cite{Indermaur2021} shown that compressive behavior of OI bone tissue is similar to healthy control. If tensile and shearing behaviors agree, fabric-strength relationships will hold too. Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and loss of homogeneity in its trabecular organization.

>>>>

\author[2,3]{Mahdi Tabatabaei}

@@

\subsection{Statistics}
The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI groups. As the groups do not have similar distributions of age and sex, a matching was performed by selecting similar individuals leading to identical mean and median age as well as identical gender distribution. For each parameter, the median value between the six ROIs from the same individual was computed. The median was preferred over the mean because it is less influenced by outliers. Normality of the distribution was assessed with QQ plot and Shapiro-Wilk test. CV had to be log-transformed to meet normal distribution assumption. Then, equal variances was assessed using Bartlett test or Brown-Forsythe test according to the normality distribution of the data. According to the normality and equal variances assumptions, t-test, Mann-Whitney test or a non-parametric permutation test was performed. The general significance level was set to 95\% for all tests. Confidence intervals were computed for t-tested variables to quantify the difference in both groups means. As Mann-Whitney tests are performed on the median, only the corresponding p-value is presented. Finally, non-parametric permutation tests are less powerful but give an empirical 95\% exclusion range and a p-value. If the difference in means belong to this exclusion range, it can be stated that group means are different with 95\% certainty.

@@

\subsection{Morphological Analysis}
The results of the morphological analysis are summarized in Table \ref{Table2}. The individual matching for age and sex allowed us to have similar group distributions with 17 females and 11 males in each group. The mean age of matched healthy individuals is 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched OI individuals. BV/TV of healthy individuals is higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01. Similarly, trabecular number is higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001. The trabecular thickness does not show significant differences between groups. On the other hand, permutation test performed for trabecular separation showed that trabecular separation is higher in mat\-ched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$). Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$). SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively. Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\

Table \ref{Table2} compares absolute values and p-values to literature. The present population age is fairly consistent with the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018}. The three other studies,  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018}, show significant differences for BV/TV, Tb.N., Tb.Sp., and Tb.Sp.SD and no significant differences for Tb.Th. In the present study, absolute values of BV/TV, TB.Th., Tb.Sp., and Tb.Sp.SD seem to be higher compared to literature. On the other hand, Tb.N. appears to be lower compared to studies in literature \cite{Folkestad2012,Kocijan2015,Rolvien2018}.

@@

\subsection{Linear Regression with Original Data Sets}
Figure \ref{02_GeneralRegression} shows the results of the linear regression analysis of each group separately, between the values of the observed stiffness tensors from \si{\micro}FE simulations and the predicted values using the Zysset-Curnier model\cite{Zysset1995} and the parameters obtained after performing the regression with linear mixed-effect model. The fitted line is represented by the dashed line, indicating the theoretical perfect correlation. For the healthy group (Figure \ref{02_Healthy}) the fit is performed on 720 ROIs leading to 8640 data points. The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%. The regression analysis for the OI group (Figure \ref{02_OI}) performed on 294 ROIs led to 3528 data points, an $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%. It can be noticed that, as the values of the observed stiffness tensor decreases, data points tend to be further apart from the diagonal (dashed line). Moreover, data the points from the stiffness tensors with lowest values are exclusively above the diagonal. The range of stiffness tensors of the OI group is wider compared to the one of the healthy group and ROIs with lower BV/TV present lower stiffness values. The values of these ROIs stiffness tensors components tends to be overestimated by the fit.\\

@@

\subsection{Filtering}
The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}. The OI data reached higher CV values and lower BV/TV values compared to healthy data. Generally, the CV tends to increase with decreasing BV/TV. The Spearman coefficient is shown above the plot as value [95\% CI]. Its value is negative and strictly different from zero. Finally, the CV threshold value used to filter the data is represented by the dashed line. It can be observed that a relatively important part of OI data will be filtered out. On the other hand, relatively few healthy data gets removed. 3D representation of extreme ROIs in terms of CV and BV/TV are shown in Appendix \ref{A2}.\\

@@

The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}. After filtering, the healthy group was reduced to 119 individuals and 603 ROIs resulting in 7236 data points, an $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\% (Figure \ref{04_Healthy}). In the OI group, more individuals were filtered leading to 38 people, 115 ROIs, and 1380 data points (Figure \ref{04_OI}). This resulted in an $R^2_{adj}$ close to 0.95 and a NE of 17\% $\pm$ 8\%.\\

@@

\subsection{BV/TV and DA Matching}
Regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}. The columns show the used data set, the fives parameters of the Zysset-Curnier model ($\lambda_0$, $\lambda_0'$, $\mu_0$, $k$, and $l$) and the assessment of fit quality ($R^2_{adj}$ and NE). Grouping healthy and OI data together for regression lead to a $k$ of 1.91 and a $l$ of 0.95. Regression result shows a $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%. The second and the last row show regression results using separated data sets and imposing the exponents $k$ and $l$ to fixed values. OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are higher than healthy one. The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively. The ANCOVA performed to quantify the group statistical significance shows a p value of 0.7.\\

@@

Table \ref{Table4} shows results obtained compared to literature. \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs. Data sets of \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} show BV/TV ranges slightly higher than in the present study and the one of \citeauthor{Gross2013} \cite{Gross2013}. On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}. Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set compared to the other studies \cite{Gross2013,Panyasantisuk2015}.\\

@@

The analysis of tBMD in relation to BV/TV is shown in Figure \ref{02_tBMD}. The t-test performed on the tBMD distributions led to a p-value <0.0001 and a 95\% CI of [-35,-13]. The linear regression performed using the linear mixed-effects model without BV/TV and group interaction provided a slope of 223 [153,292] (value [95\% CI]). The intercept value was 534 [518,550] and the group variable led to a value of 14 [7,20]. The prediction using the fixed effects only led to a $R^2$ of 0.23 and a SE of 33. Then, using the linear mixed-effect model with the interaction regressor (BV/TV x Group), this added variable presented a value of 3 [-66,72] and a p-value of 0.94.

@@

As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare morphological parameters. The imaging system explains most of the differences between the absolute morphological values of the present study compared to the others. \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed their measurements on first generation XCT scanners with a voxel size of 82 \si{\micro}m, while we have used a second generation XCT with a voxel size of 61 \si{\micro}m. The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types. They showed that BV/TV, Tb.Th., and Tb.Sp. are higher in second generation XCT scanners and Tb.N. is lower compared to first generation XCT. These results give confidence in our observed values. Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length. This conditions the Tb.N. and Tb.Sp. as they depend on the ROI size. Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain a portion of trabecular bone. Even with the low sample size (2x28 individuals), the statistic tests have shown significant differences between groups with the more significant being for the CV values. The CV values show that heterogeneity of OI trabecular bone is higher compared to healthy control and the more discriminant parameter. Finally, the significant differences observed in BV/TV and DA even with matched age and gender justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the fit must be performed on identical ranges to obtain comparable values.\\

The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i. e. slightly lower than \citeauthor{Gross2013}\cite{Gross2013} and \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}) for the healthy group. Components of the stiffness tensors are distributed to both sides of the diagonal. On the other hand, the linear regressions performed in this study using the OI original data set presents lower $R^2_{adj}$ and higher NE than such fit reach usually \cite{Gross2013,Panyasantisuk2015}. The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation. These differences come from ROIs presenting a low stiffness. The regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to $10^0$ MPa and lower, the fit tends to overestimate the stiffness. This is because ROI stiffness is highly dependent on BV/TV values. Some ROIs with low BV/TV don't have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}. Trying to homogenize such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}). Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}. An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.\\

Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV. It shows a tendency of CV values to increase with decreasing BV/TV values. Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly become highly heterogeneous. A simple assumption about this relation is that it could be monotonic. Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation. As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenization seem feasible. However, the value of this threshold actually results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\

The linear regressions performed on filtered data sets present direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values. For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) compared to the linear regression using the unfiltered data set is negligible. On the other hand, NE values are decreased by 2\% and therefore improved. The filtering eliminates data points further away from the diagonal (better NE) and other data points close to the diagonal leading to a smaller number of points (modifying $R^2_{adj}$). For the OI group, filtering leads to an important improvement of the linear regression (i.e. higher $R^2_{adj}$ and lower NE). $R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group. These results give confidence to the filtering procedure and are a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\

After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets. This allows one to determine values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m. Imposing these values to perform the linear regression on data sets of the matched individuals allows us to highlight differences, if any, between healthy and OI trabecular bone. The relatively low differences for $\lambda_0'$ and $\mu_0$ once again supports the hypothesis for similar fabric-elasticity relationships between healthy and OI trabecular bone. For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but the 95\% CI intervals still show a common range which almost include both the $\lambda_0$ of OI and healthy linear regressions. Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group showed a p-value far above the 5\% significance level. With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone. In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity. Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\

Imposing $k$ and $l$ allows one to estimate the effect of different image resolutions.  \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened them to 36 \si{\micro}m. \citeauthor{Gross2013}\cite{Gross2013} showed that different anatomical locations lead to only slight differences. Comparing regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution. Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones. Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap. As for the filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies. Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants as a result of the different image resolutions. The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this increase of DA. Lower stiffness constants are obtained because the coarser structure resulting from XCTII can't be as optimized as the fine detailed structure obtained by \si{\micro}CT. Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimized morphology of trabecular bone with a high fidelity. By decreasing the scan spatial resolution, the scanned structure becomes bulkier. Performing a linear regression on this less optimized structure leads to the observed lower stiffness constants. Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality. Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant. On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.\\

@@

The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans. Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics. Effectively as those patients are in wheelchair, it could be interesting to analyze the impact of this condition on the weight bearing tibia. The ROI homogeneity has an important impact on the analysis quality. As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting. More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models could be negligible if the proportion of such low stiffness ROIs stays low. A similar study could be performed on XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution. Another limitation is that the scans were performed on different devices which were not cross-calibrated. However, as they are the same model, it is expected to have a minor impact.\\

In conclusions, the samples analyzed in the present study had similar morphology compared to data reported in the literature. We couldn't find differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e. with a CV lower than 0.263. \citeauthor{Indermaur2021}\cite{Indermaur2021} could show that the compressive behaviour of OI bone tissue is similar to the one of healthy control at the ECM level. If the tensile and shearing behaviour is similar as well, fabric-strength relationships will hold too. Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and the loss of homogeneity in its trabecular organization.
